---
title: "Homework 3 - Time Series ETS"
author: "Prof. Dr. Stephan Trahasch"
date: 'Submission date: 28.01.2021'
params:
  datadir: ../data/
output:
  word_document:
    toc: yes
    toc_depth: '3'
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp2)
library(fpp3)
```

# Exercise 1 - Sales of paperback books
Data set `books` contains the daily sales of paperback and hardcover books at the same store. The task is to forecast the next four days sales for paperback and hardcover books.

Plot the series and discuss the main features of the data.

```{r}
# todo
autoplot(books)
```

Your answer:

Beide Zeitreihen zeigen einen positiven Trend. Sie haben einen ähnlichen Verlauf. Es lässt sich keine Zyklität oder Saisonalität erkennen. 

Both series are trended, with a similar slope. There is no obvious weekly seasonality.


```{r}
hardcover <- books %>% as_tsibble(pivot_longer = FALSE) %>% select(Hardcover)
paperback <- books %>% as_tsibble(pivot_longer = FALSE) %>% select(Paperback)
```

Compare the forecasting performance of the  exponential smoothing methods SES and Holt to the paperback series.

```{r}
# todo
fit <- paperback %>%
  model(
    ses = ETS(Paperback ~ error("A") + trend("N") + season("N")),
    holt = ETS(Paperback ~ error("A") + trend("A") + season("N"))
    )
report(fit)
```


```{r}
# Forecast 4 days
# todo

fc = fit %>% forecast(h=4)
```
  
Plot the 4 day forecasts for hardcover.

```{r}
# todo 

autoplot(paperback) + autolayer(fc)

```

Discuss the merits of the two forecasting methods. 
Compare the forecasts for the series using both methods. Which do you think is best?


```{r}
# to do 
# calculate accuracy of training data
accuracy(fit)
```
Your answer:
Das "holt" Modell liefert leicht bessere Metriken. Der RMSE und MAE Wert sind deutlicht besser. Der MAPE Wert ist nahezu identisch.
Der forecast des "holt" Modell ist besser geeignet. Er bezieht den Trend der Daten mit ein. Beide Modelle beziehen die vorherigen Daten gut ein.


```{r}
# plot components of model
# todo
components(fit) %>% autoplot()

```
