---
title: "Homework 3 - Time Series ARIMA"
author: "Prof. Dr. Stephan Trahasch"
date: 'Submission date: 28.01.2021'
params:
  datadir: ../data/
output:
  word_document:
    toc: yes
    toc_depth: '3'
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp2)
library(fpp3)
```

# Exercise 1 - International Visitors to Australia
Consider `austa`, the total international visitors to Australia (in millions) for the period 1980-2015.

Use arima to find an appropriate auto ARIMA model. 
What model was selected? 
Check that the residuals look like white noise. 
Plot forecasts for the next 10 periods.

```{r}
# todo
austa_ts <- as_tsibble(austa)
autoplot(austa_ts)
fit_1 <- austa_ts %>% model(
  auto = ARIMA(value)
  )  
report(fit_1)
# ARIMA(0,1,1)
```

Check if residuals look like white noise.

```{r}
# todo
gg_tsresiduals(fit_1)
# your answer
# Die residuals sehen wie white noise aus
```

Forecast for 10 years and plot the forecast

```{r}
# todo

fc_1 = fit_1 %>% forecast(h=10)
autoplot(austa_ts) + autolayer(fc_1) 

```

Now add following ARIMA models 
* ARIMA(0,1,1) with no constant 
  (you can include a constant using `ARIMA(y~1)` to the formula or exclude it by adding `ARIMA(y~0)`)
* ARIMA(0,1,0) (removing MA term)
* ARIMA with stepwise false 

```{r}
# todo 
fit_2 <- austa_ts %>% model(
  arima_011 = ARIMA(value ~ 0 + pdq(0,1,1)),
  arima_010 = ARIMA(value ~ pdq(0,1,0)),
  arima_sw = ARIMA(value, stepwise = FALSE)
  )  
```


```{r}
report(fit_2)
fit_2 %>%
  glance() %>%
  arrange(AICc)

```
What is the difference?

Your answer:

Der AICc Wert für das arima_sw ist am niedrigsten. arima_010 liegt sehr nah am arima_sw.Das arima_011 Modell hat den schlechtesten AICc Wert.


Plot forecasts from an ARIMA(2,1,3) model with drift with constant.

```{r}
# todo
fit_3 <- austa_ts %>% model(
  arima213 = ARIMA(value ~ 1 + pdq(2,1,3)),
  )  
report(fit_3)
```

```{r}
fit_3 %>% forecast %>% autoplot(austa)
```

Plot forecasts from an ARIMA(2,1,3) model with drift and with NO constant.
What happens?

```{r}
# todo
fit_4 <- austa_ts %>% model(
  arima213 = ARIMA(value ~ 0 + pdq(2,1,3)),
  )  
report(fit_4)
```

```{r}
# 1 error encountered for arima213
# [1] non-stationary AR part from CSS
# The model fitting causes an error because the drift is really needed. What happens is that the AR terms 
# try to handle the drift but end up going outside the stationarity region.

```

Report and plot forecasts from an ARIMA(0,0,1) model with a constant.

```{r}
# todo 
fit_5 <- austa_ts %>% model(
  arima001 = ARIMA(value ~ 1 + pdq(0,0,1)),
  ) 
report(fit_5)
```

```{r}
fit_5 %>% forecast %>% autoplot(austa)
```


Remove the constant and plot again.

```{r}
# todo 
fit_6 <- austa_ts %>% model(
  arima001 = ARIMA(value ~ 0 + pdq(0,0,1)),
  ) 
report(fit_6)
```

```{r}
fit_6 %>% forecast %>% autoplot(austa)
```

What is the difference?

Your answer:

Die Prediction Intervalle sind für das Model ohne Konstante viel größer. Die prädiktierten Werte fallen ohne Konstante ebenfalls stärker.
Beide Forecasts sehen nicht schlüssig aus.


Plot forecasts from an ARIMA(0,2,1) model with no constant.

```{r}
# todo
fit_7 <- austa_ts %>% model(
  arima021 = ARIMA(value ~ 0 + pdq(0,2,1)),
  ) 
report(fit_7)
```

```{r}
fit_7 %>% forecast %>% autoplot(austa)

# The second level of differencing induces a linear trend in the forecasts.
```


# Exercise 2 - Quarterly US GDP

In this exercise we use the `usgdp` series. If necessary, find a suitable Box-Cox transformation for the data.

```{r}
autoplot(usgdp)
```

```{r}
autoplot(BoxCox(usgdp, lambda = 'auto'))
lambda <- BoxCox.lambda(usgdp)
lambda
# BoxCox tranformation nicht benötigt. Varianz muss nicht stabilisiert werden.
```

```{r}
usgdp_ts <- as_tsibble(usgdp)
```

Fit a suitable ARIMA model to the transformed data using auto ARIMA;

```{r}
# todo
fit_8 <- usgdp_ts %>% model(
  auto = ARIMA(value)
  )  
report(fit_8)
```

Try some other plausible models by experimenting with the orders chosen.


```{r}
# todo
# The second order differencing will induce a trend in the forecasts, which is required here, so we will look at changing only $p$ and $q$.

fit_9 <- usgdp_ts %>% model(
  fit020 = ARIMA(value ~ pdq(0,2,0)),
  fit021 = ARIMA(value ~ pdq(0,2,1)),
  fit022 = ARIMA(value ~ pdq(0,2,2)),
  fit023 = ARIMA(value ~ pdq(0,2,3)),
  fit120 = ARIMA(value ~ pdq(1,2,0)),
  fit121 = ARIMA(value ~ pdq(1,2,1)),
  fit122 = ARIMA(value ~ pdq(1,2,2)),
  fit123 = ARIMA(value ~ pdq(1,2,3)),
  fit220 = ARIMA(value ~ pdq(2,2,0)),
  fit221 = ARIMA(value ~ pdq(2,2,1)),
  fit222 = ARIMA(value ~ pdq(2,2,2)),
  fit223 = ARIMA(value ~ pdq(2,2,3)),
  fit320 = ARIMA(value ~ pdq(3,2,0)),
  fit321 = ARIMA(value ~ pdq(3,2,1)),
  fit322 = ARIMA(value ~ pdq(3,2,2)),
  fit323 = ARIMA(value ~ pdq(3,2,3)),
  ) 


```
```{r}
# todo
report(fit_9)

```
Choose what you think is the best model using AICc (use which.min(c(yourmodel1$aicc,...) and check the residual diagnostics;
Change eval=TRUE!

```{r}
# fitxxx are the fitted models
fit_9 %>%
  glance() %>%
  arrange(AICc)
```
```{r}
# todo 
# check residuals of best model
# The best according to the AICc values is the ARIMA(2,2,1) model.

fit_9 %>% select(fit221) %>% gg_tsresiduals()
#fit222 Das acf plot zeigt einen Ausschlag in lag 12. Das Histogram entspricht nicht ganz einer Normalverteilung.
#The residuals pass the Ljung-Box test, but the histogram looks like it has heavier tails than Gaussian.
```
```{r}
augment(fit_9) %>% features(.innov, ljung_box, lag=10, dof=4) %>% arrange(lb_pvalue)
# fit221 besteht den ljung_box test
```


Produce forecasts of your fitted model. Do the forecasts look reasonable?

```{r}
# todo
fit_9 %>% select(fit221) %>% forecast(h=20) %>% autoplot(usgdp)
# These look reasonable.
```

Compare the results with what you would obtain using `ets()` (with no transformation).

```{r}
# todo

fit_10 <- ets(usgdp)
fc <- forecast(fit_10, h=20) 
autoplot(fc) 

# What is the difference?
# Die Prediction Intervalle sind für das Arima Model kleiner. Die Steigung der Prediction von ARIMA flacht im Gegensatz zum ets model ab.
```

